BIP Architectures and Open pNets

### 2019-02-01 2018-Arch-pNets/arch-pNets.tex
skype changes and shorter 2.1

%****************************************************************
We assume the existence of a term algebra $\Talg$,
where $\signature$ is the signature of the data and action constructors,
and $\variables$ a set of \emph{variables}. Within $\Talg$, we distinguish a set of
\emph{data expressions} $\exprs{\variables}$; $e_i$ ranges over boolean expressions, and a set of \emph{Boolean
expressions} 
$\boolexprs{\variables}\subseteq\exprs{\variables}$, $g$ (guards) ranges over boolean expressions.
On top of $\exprs{\variables}$ we build the \emph{action algebra}
$\actions{\variables}$, with $\actions{\variables}\subseteq\Talg$.
We  use
\[
\assigns{\variables} \bydef{=}
\Setdef{(x_i := e_i)^{i \in I}}{
  x_i^{i \in I}\subseteq \variables,
  e_i^{i \in I}\subseteq \exprs{\variables},
  I \text{ is a finite index set}
}
\]
to denote the set of \emph{variable assignments}.  $u$ ranges over sets of assignments.
To be able to reason about the data flow between actions, we
distinguish \emph{input variables} of the form $?x$ within terms; the function
$\vars(t)$ identifies the set of variables in a term
$t\in\Talg$, and $iv(t)$ returns its input variables.

@@

\paragraph{pNets definition.}
\label{section:pNets}
A pLTS is a labelled transition system with variables that can be
used to characterise a state, as part of actions, guards, and
assignments. 
%Without loss of generality and to simplify the formalisation, we suppose 
%here that 
%variables are local to each 
%state: each state has its set of variables disjoint from the others. Transmitting 
%variable values from one state to the other can be done by explicit assignment. 
%Similarly, to simplify the management of variables and without loss of expressivity, we 
%suppose that transitions looping to the same state does not do assignments.
%Note that we make no assumption on the finiteness of the set of states, nor
%on the finite branching of the transition relation.
We  define the set of actions a pLTS can use ($a$
ranges over action labels):\\
%. Action terms are  of the form 
$\alpha=a(?x_i^{i \in I}, 
\Expr_j^{j \in J})$, where $?x_i^{i \in I}$ are input variables, $\Expr_j^{j \in J}$ 
are expressions.

@@

\begin{example}[An open-transition]
  \label{Example:TimerOT3}
  The \texttt{Failure TimerMonitor} pNet of \Fig{FailureTimer:pNet} has \replaceSB{2 controllers 
  and 2 holes. O}{two coordinators and four dangling ports, which can be grouped into two holes (see \secn{encoding} below).  Disregarding the maximal progress priority, o}ne of its possible open-transitions is:\\[-2ex]
%%   \footnote{Note the underlined action notation that we use to represent synchronised actions, i.e. actions that are the result of a synchronisation and that we declare as \emph{non-observable}. Those actions are only used for monitoring purposes and statement of properties but cannot be synchronised again in a synchronisation vector. This allows us to define a weak bisimilarity equivalence with p=compositional properties.}
%%   : 
%% \noteLH{underline notation might be a bit too much for this paper : see at the end if we remove it or not (will we use it?)}
% \smallskip
%$  OT_1  = \openrule
%{\emptyset, \{t:=0, z:=TopInt, zone:=[minZone,maxZone]\} }
%{\ostate{00} \OTarrow{\underline{\alpha_4}} \ostate{11}}
%$
%\medskip
\begin{small}\[ \openrule
  {\{B\mapsto fail(b0)\},\\
    t<GetMax(z) \land b1=true \land b1=>b0 \land z'= (z \cap (b1?zone;TopInt)+t) {\color{blue}\land (t' = t) \land (zone' = zone)},\\
    \{z:=z', {\color{blue}t := t'+1, zone := zone'}\} }
  {\ostate{11} \OTarrow{\alpha_0} \ostate{12}}
  \]
\end{small}\vspace{-1.5ex}
%
%The first one, $OT_1$, corresponds to the simultaneous initialisation
%of both controlers. No hole is active at that time, and the resulting
%global action is denoted $\alpha_4$.

>>>>

%****************************************************************
We assume the existence of a term algebra $\Talg$,
where $\signature$ is the signature of the data and action constructors,
and $\variables$ a set of \emph{variables}. Within $\Talg$, we distinguish a set of
\emph{data expressions} $\exprs{\variables}$, $e$ ranges over  expressions; and a set of \emph{Boolean
expressions} 
$\boolexprs{\variables}\subseteq\exprs{\variables}$, $g$ (guards) ranges over boolean expressions.
On top of $\exprs{\variables}$ we build the \emph{action algebra}
$\actions{\variables}$, with $\actions{\variables}\subseteq\Talg$.
We define $\assigns{\variables}$ as the set of variable assignments of the form: $(x_i := e_i)^{i \in I}$. 
%\[
%\assigns{\variables} \bydef{=}
%\Setdef{(x_i := e_i)^{i \in I}}{
%  x_i^{i \in I}\subseteq \variables,
%  e_i^{i \in I}\subseteq \exprs{\variables},
%  I \text{ is a finite index set}
%}
%\]
%to denote the set of \emph{variable assignments}. 
 $u$ ranges over sets of assignments.
%To be able to reason about the data flow between actions, we
%distinguish \emph{input variables} of the form $?x$ within terms; t
The function
$\vars(t)$ identifies the set of variables in a term.
%$t\in\Talg$, and $iv(t)$ returns its input variables.

@@

\paragraph{pNets definition.}
\label{section:pNets}
A pLTS is a labelled transition system with variables that can be
used to characterise a state, as part of actions, guards, and
assignments. 
%Without loss of generality and to simplify the formalisation, we suppose 
%here that 
%variables are local to each 
%state: each state has its set of variables disjoint from the others. Transmitting 
%variable values from one state to the other can be done by explicit assignment. 
%Similarly, to simplify the management of variables and without loss of expressivity, we 
%suppose that transitions looping to the same state does not do assignments.
%Note that we make no assumption on the finiteness of the set of states, nor
%on the finite branching of the transition relation.
We need to introduce input variables to reason about the data flow between actions.
We  define the set of actions a pLTS can use ($a$
ranges over action labels):\\
% Action terms are  of the form 
$\alpha=a(?x_i^{i \in I}, 
\Expr_j^{j \in J})$, where $?x_i^{i \in I}$ are input variables, $\Expr_j^{j \in J}$ 
are expressions.

@@

\begin{example}[An open-transition]
  \label{Example:TimerOT3}
  The \texttt{Failure TimerMonitor} pNet of \Fig{FailureTimer:pNet} has \replaceSB{2 controllers 
  and 2 holes. O}{two coordinators and four dangling ports, which can be grouped into two holes (see \secn{encoding} below).  Disregarding the maximal progress priority, o}ne of its possible open-transitions is:\\[-2ex] \noteLH{maximal progress undefined}
%%   \footnote{Note the underlined action notation that we use to represent synchronised actions, i.e. actions that are the result of a synchronisation and that we declare as \emph{non-observable}. Those actions are only used for monitoring purposes and statement of properties but cannot be synchronised again in a synchronisation vector. This allows us to define a weak bisimilarity equivalence with p=compositional properties.}
%%   : 
%% \noteLH{underline notation might be a bit too much for this paper : see at the end if we remove it or not (will we use it?)}
% \smallskip
%$  OT_1  = \openrule
%{\emptyset, \{t:=0, z:=TopInt, zone:=[minZone,maxZone]\} }
%{\ostate{00} \OTarrow{\underline{\alpha_4}} \ostate{11}}
%$
%\medskip
\begin{small}\[ \openrule
  {\{B\mapsto fail(b0)\},\\
    t<GetMax(z) \land b1=true \land b1=>b0 \land z'= (z \cap (b1?zone;TopInt)+t) {\color{blue}\land (t' = t) \land (zone' = zone)},\\
    \{z:=z', {\color{blue}t := t'+1, zone := zone'}\} }
  {\ostate{11} \OTarrow{\alpha_0} \ostate{12}}
  \]
\end{small}\vspace{-1.5ex}
%
%The first one, $OT_1$, corresponds to the simultaneous initialisation
%of both controlers. No hole is active at that time, and the resulting
%global action is denoted $\alpha_4$.
